---
title: "PA1_template.Rmd"
author: "Scott Shander"
date: "Tuesday, October 07, 2014"
output: html_document
---

Loading and preprocessing the data
----------------
Let's read the file in and format the date varlable. Also let's create another variable for day, to use in aggregating totals and averages.

```{r, echo=TRUE}
setwd("C:/Users/Owner/Documents/R")
act<-read.csv("activity.csv")
act$date <-as.Date(act$date)
act$day <- substring(act$date,9,10)
```

What is mean total number of steps taken per day?
----------------
below is a histogram, along with the mean and median.
```{r, echo=TRUE}
total_steps_per_day <-tapply(act$steps,act$day, sum)
hist(total_steps_per_day)
mean <- mean(total_steps_per_day,na.rm=TRUE)
median <- median(total_steps_per_day,na.rm=TRUE)
mean
median

```

What is the average daily activity pattern?
--------------------------
Notice the chart of the average activity over time throughout the day.
```{r, echo=TRUE}
avg_steps<- aggregate(act$steps, list(interval=act$interval), mean, na.rm=TRUE)
plot(avg_steps$interval,avg_steps$x,type = "l")
avg_steps$interval[which.max(avg_steps$x)]
```
Imputing missing values
-------------------------
here, I calculated the average for each 5-minute interval and imputed for all missing values.  The change as has a marginal impact on the calculations. 
```{r, echo=TRUE}
sum(is.na(act))
library(plyr)
act2 <- join(act, avg_steps, by = "interval")
act2$all_steps <-ifelse(!is.na(act2$steps),act2$steps,act2$x)
act2 <- act2[,c(6,2,3,4)]
total_steps_per_day_fill <-tapply(act2$all_steps,act2$day, sum)
hist(total_steps_per_day_fill)
mean_fill <- mean(total_steps_per_day_fill,na.rm=TRUE)
median_fill <- median(total_steps_per_day_fill,na.rm=TRUE)
mean_fill
median_fill
mean_fill ==mean
median_fill==median
mean_fill-mean
```
Are there differences in activity patterns between weekdays and weekends?
-----------------
It appears so when you look at the chart!
```{r, echo=TRUE}
act2$weekday <- as.factor(ifelse(weekdays(act2$date)=="Saturday"|weekdays(act2$date)=="Sunday","weekend","weekday"))
act3 <- act2[act2$weekday == "weekday",]
act4 <- act2[act2$weekday =="weekend",]
avg_steps_weekday<- aggregate(act3$all_steps, list(interval=act3$interval), mean)
avg_steps_weekend<- aggregate(act4$all_steps, list(interval=act4$interval), mean)
plot(avg_steps_weekday$interval,avg_steps_weekday$x,type = "l")
lines(avg_steps_weekend$interval,avg_steps_weekend$x,type = "l", col="red")
legend(cex=0.7,"topright",c("weekday","weekend"), col=c("black","red"), pch=1, pt.cex = 1)
```